Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-07-27 paper.tex
clean
Small updates in outline
Restored files permissions
Removed figure pdfs. Added all gnuplot scripts to generate_figures.sh.

The buffer might be a
slightly complex data structure, for instance an associative array or
a Python dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Figure~\ref{fig:cluster-reads-buffer}
illustrates how the buffer would fill up for the two first blocks in a reconstructed image.Implementation details about the
buffer are given later.
\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/buffer.pdf_tex}
\caption{Illustration of the buffer used in cluster reads, showing how
  the two first 4x4 blocks of a reconstructed image would be arranged
  in the buffer.  White portions in the buffer are not allocated.  The
  column-major order is used, similar to the NIfTI format: i is the
  fastest changing dimension, followed by j, followed by k.}
\label{fig:cluster-reads-buffer}
% Data file: path
% Gnuplot file: path
% Script file: path
\end{figure}

@@

\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/case1-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case2-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case3.pdf_tex}
\caption{Memory-load configurations in cluster reads, leading to
  different number of seeks. Red blocks need seeking before each of
  their rows ($d^2$ seeks). Blue blocks need seeking before each of
  their slices ($d$ seeks in total). Green blocks need only a single
  seek. Grey, dashed, transparent blocks represent the contiguous
  memory loads and are added for the sake of visualization.}
\label{fig:cluster-reads}
\end{figure}
Our cluster reads algorithm focuses on the three memory load
configurations represented in Figure~\ref{fig:cluster-reads}, that is,
the amount of memory $m'$ used by the algorithm is rounded down to the
closest number of complete blocks (case 1), complete block rows (case
2) or complete block slices (case 3). This is in general reasonable
since adding an incomplete row to a set of complete ones multiplies
the number of required seeks by $d$, as illustrated in
Figure~\ref{fig:avoided-configurations}-Left. In some cases though,
rounding $m$ down to $m'$ might increase the number of required
memory load to a point that the overall number of seeks also
increases. However, such cases are a bit pathological and their
complete description requires extensive calculations involving modulo
arithmetic, which we felt were unwieldy to report here.
\begin{figure}
  \centering
\def\svgwidth{0.3\columnwidth}
\input{figures/incomplete-rows.pdf_tex}
\hfill
\def\svgwidth{0.3\columnwidth}
\input{figures/overlap.pdf_tex}
\caption{Configurations that increase the number of seeks per
  memory-load and are thus deliberately avoided by cluster
  reads. Left: configuration with incomplete block rows (multiplies
  number of seeks by $d$). Right: configuration with block rows that
  overlap multiple block slices (multiplies number of seeks by $2$).}
\label{fig:avoided-configurations}
\end{figure}
To summarize, the amount of memory used $m'$ is set as follows in each of the 3 cases:
%% \begin{equation*}
%%   m_{case 1}' = bd^3\floor*{\frac{m}{d^3b}}; m_{case 2}' = bd^3\sqrt[3]{n}\floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = bd^3\sqrt[3]{n}^2\floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
%% \end{equation*}
%% which gives:
\begin{equation*}
  m_1' = \frac{Rb}{n}\floor*{\frac{mn}{Rb}};
  m_2' = \frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
  m_3' = \frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}
\end{equation*}

@@

@@

\subsubsection{Implementation}

\subsubsection{Data}
To benchmark our library we use the 76~GB BigBrain image split in 125
non-overlapping chunks. The data is available \todo{[here]}.

\subsubsection{Infrastructure}

@@

>>>>

The buffer might be a
slightly complex data structure, for instance an associative array or
a Python dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Figure~\ref{fig:cluster-reads-buffer}
illustrates how the buffer would fill up for the two first blocks in a reconstructed image.Implementation details about the
buffer are given later.
\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/buffer.pdf_tex}
\caption{Illustration of the buffer used in cluster reads, showing how
  the two first 4x4 blocks of a reconstructed image would be arranged
  in the buffer.  White portions in the buffer are not allocated.  The
  column-major order is used, similar to the NIfTI format: i is the
  fastest changing dimension, followed by j, followed by k.}
\label{fig:cluster-reads-buffer}
% Data file: path
% Gnuplot file: path
% Script file: path
\end{figure}

@@

\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/case1-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/case2-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/case3.pdf_tex}
\caption{Memory-load configurations in cluster reads, leading to
  different number of seeks. Red blocks need seeking before each of
  their rows ($d^2$ seeks). Blue blocks need seeking before each of
  their slices ($d$ seeks in total). Green blocks need only a single
  seek. Grey, dashed, transparent blocks represent the contiguous
  memory loads and are added for the sake of visualization.}
\label{fig:cluster-reads}
\end{figure}
Our cluster reads algorithm focuses on the three memory load
configurations represented in Figure~\ref{fig:cluster-reads}, that is,
the amount of memory $m'$ used by the algorithm is rounded down to the
closest number of complete blocks (case 1), complete block rows (case
2) or complete block slices (case 3). This is in general reasonable
since adding an incomplete row to a set of complete ones multiplies
the number of required seeks by $d$, as illustrated in
Figure~\ref{fig:avoided-configurations}-Left. In some cases though,
rounding $m$ down to $m'$ might increase the number of required
memory load to a point that the overall number of seeks also
increases. However, such cases are a bit pathological and their
complete description requires extensive calculations involving modulo
arithmetic, which we felt were unwieldy to report here.
\begin{figure}
  \centering
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/incomplete-rows.pdf_tex}
\hfill
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/overlap.pdf_tex}
\caption{Configurations that increase the number of seeks per
  memory-load and are thus deliberately avoided by cluster
  reads. Left: configuration with incomplete block rows (multiplies
  number of seeks by $d$). Right: configuration with block rows that
  overlap multiple block slices (multiplies number of seeks by $2$).}
\label{fig:avoided-configurations}
\end{figure}
To summarize, the amount of memory used $m'$ is set as follows in each of the 3 cases:
%% \begin{equation*}
%%   m_{case 1}' = bd^3\floor*{\frac{m}{d^3b}}; m_{case 2}' = bd^3\sqrt[3]{n}\floor*{\frac{m}{d^3b\sqrt[3]{n}}}; m_{case 3}' = bd^3\sqrt[3]{n}^2\floor*{\frac{m}{d^3b\sqrt[3]{n}^2}}
%% \end{equation*}
%% which gives:
\begin{equation*}
  m_1' = \frac{Rb}{n}\floor*{\frac{mn}{Rb}};
  m_2' = \frac{Rb}{\sqrt[3]{n}^2}\floor*{\frac{m\sqrt[3]{n}^2}{Rb}};
  m_3' = \frac{Rb}{\sqrt[3]{n}}\floor*{\frac{m\sqrt[3]{n}}{Rb}}
\end{equation*}

@@

\subsection{Implementation}

@@

\subsection{Data}
To benchmark our library we use the 76~GB BigBrain image split in 125
non-overlapping chunks. The data is available \todo{[here]}.

\subsection{Infrastructure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-ssd.pdf}
  \hfill
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-comparision-hdd.pdf}\\
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
  \hfill
    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
  \caption{Comparison between Multiple reads, blocks and slices. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
\label{fig:multiple-reads}
\end{figure}
